Central Limit Theorems for Super Ornstein-Uhlenbeck Processes with stable Branching

### 2019-08-23 supCLT.tex
change supCLT.tex
Stage the file
change supCLT.tex


change supCLT.tex


change supCLT.tex tep


change supCLT.tex

@@

% ** Motivation

@@

\label{sec: Motivation}

@@

  - \alpha z + \rho z^2 + \int_{(0,\infty)} (e^{-zy} - 1 + zy) \pi(dy)

@@

where $\alpha > 0 $, $\rho \geq0$ and $\pi$ is a measure on $(0,\infty)$ with $\int_{(0,\infty)}(y\wedge y^2) \pi(dy)< \infty$.
$\psi$ is referred to as a branching mechanism and $\pi$ is referred to as the L\'evy measure of $\psi$.

@@

For each $f,g\in \mathcal B(\mathbb R^d, \mathbb R)$ and $\mu \in \mathcal M(\mathbb R^d)$, write $\langle f,\mu\rangle = \int f(x)\mu(dx)$ and $\langle f, g\rangle = \int f(x)g(x) dx$ whenever the integrals make sense.

@@

where $(t,x) \mapsto V_tf(x)$ is the unique locally bounded positive solution to the equation

@@

The strong law of large numbers for the super-OU process $X$ can be stated as follows:
Under some conditions on the branching mechanism $\psi$ (these conditions are satisfied under our Assumptions 1 and 2 below), there exists an $\Omega_0$ of $\mathbb{P}_\mu$-full probability for every $\mu\in\mathcal M(\mathbb R^d)$ such that on $\Omega_0$, for every Lebesgue-a.\/e. continuous bounded non-negative function $f$ on $\mathbb R^d$, we have $\lim_{t\to\infty} e^{-\alpha t} \langle f, X_t\rangle =H_\infty\langle f, \varphi\rangle $, where $H_\infty$ is the limit of the martingale $e^{-\alpha t}\langle 1,X_t\rangle$ and $\varphi$ is the invariant density of the OU process $\xi$ defined in \eqref{invariantdensity} below.
See \cite[Theorem 2.13 \& Example 8.1]{ChenRenYang2019Skeleton} and \cite[Theorem 1.2 \& Example 4.1]{EckhoffKyprianouWinkel2015Spines}.

In this paper, we will establish some spatial central limit theorems (CLT) for the super-OU process $X$ above.
Our key assumption is that the branching mechanism $\psi(z)$ takes the form of $-\alpha z +\rho z^2+ \psi_1(z)$, where $\psi_1$ is close to $\eta z^{1+\beta}$ with $\eta>0$ and $\beta\in (0, 1)$, in some sense.
We want to find $(F_t)_{t\geq 0}$ and $(G_t)_{t\geq 0}$ such that $ (\langle f, X_t \rangle -G_t)/F_t $ converges weakly to some non-degenerate random variable as $t\rightarrow\infty$, for a large class of functions $f$.
It turns out that the statements of the CLT's are different in three regimes depending on the sign of $\alpha\beta-\kappa_f b (1+\beta)$: 
the small branching rate regime $\alpha\beta< \kappa_f b(1+\beta)$; 
the critical branching rate regime $\alpha\beta = \kappa_f b(1+\beta)$; 
and the large branching rate regime $\alpha\beta > \kappa_f b(1+\beta)$. 
Here, $\kappa_f$ is the order of $f$ in the spectral decomposition in terms of the spectrum of $L$.
Note that, in the setting of this paper, $\langle f,X_t\rangle$ typically have infinite second moment.

There are many papers studying central limit theorems for branching processes, branching diffusions and superprocesses under the second moment condition.
See \cite{Heyde1970A-rate, HeydeBrown1871An-invariance, HeydeLeslie1971Improved} for supercritical Galton-Watson processes (GW processes), \cite{KestenStigum1966Additional,KestenStigum1966A-limit} for supercritical multi-type GW processes \cite{Athreya1969Limit,Athreya1969LimitB,Athreya1971Some} for supercritical multi-type continuous time branching processes and \cite{AsmussenHering1983Branching} for general supercritical branching Markov processes under certain conditions.
Some spatial central limit theorems for   supercritical branching OU processes with binary branching mechanism were proved in \cite{AdamczakMilos2015CLT} and some spatial central limit theorems for supercritical super-OU processes with branching mechanisms satisfying a fourth moment condition were proved in \cite{Milos2012Spatial}.
These two papers made connections between the CLT and  the branching rate regimes.
Some spatial central limit theorems for supercritical super-OU  processes with branching mechanisms satisfying only a second moment condition were established in \cite{RenSongZhang2014Central}.
Moreover, compared with the results of \cite{AdamczakMilos2015CLT,Milos2012Spatial}, the limit distributions in \cite{RenSongZhang2014Central} are non-degenerate.
Since then, a series of spatial central limit theorems for a large class of general supercritical branching Markov processes and superprocesses with spatially dependent branching mechanisms were proved in \cite{RenSongZhang2014CentralB,RenSongZhang2015Central,RenSongZhang2017Central}.
The functional version of the central limit theorems was established in \cite{Janson2004Functional} for supercritical multitype branching processes, and  in \cite{RenSongZhang2017Functional} for supercritical superprocesses.

@@

% ** Main results
\subsection{Main results}
\label{sec: main results}{}
We will always assume that the following assumption holds.
\begin{asp}
  \label{asp: Greys condition}
  The branching mechanism satisfies Grey's condition, i.e., there exists $z' > 0$ such that $\psi(z) > 0$ for all $z>z'$ and  $\int_{z'}^\infty \psi(z)^{-1}dz < \infty$.
\end{asp}
For each $\mu \in \mathcal M(\mathbb R^d)$, write $\|\mu\| = \langle 1, \mu\rangle$.
It is known (see \cite[Theorems 12.5 \& 12.7]{Kyprianou2014Fluctuations} for example) that, under Assumption \ref{asp: Greys condition}, the \emph{extinction event} $D :=\{\exists t\geq 0,~\text{s.t.}~ \|X_t\| =0 \}$ has positive probability, with respect to $\mathbb P_\mu$ for each  $\mu \in \mathcal M(\mathbb R^d)$.
In fact, $ \mathbb{P}_{\mu} (D) = e^{-\bar v \|\mu\|}, $ for $\mu\in \mathcal M(\mathbb R^d)$, where $ \bar v := \sup\{\lambda \geq 0: \psi(\lambda) = 0\} \in (0,\infty) $ is the largest root of $\psi$.

Denote by $\Gamma$ the gamma function.
For any $\sigma$-finite signed measure $\mu$, we use $|\mu|$ to denote the total variation measure of $\mu$.
In this paper, we will also assume the following:
\begin{asp}
  \label{asp: branching mechanism}
  There exist constants $\eta > 0$ and $\beta \in (0,1)$ such that
  \begin{align}
    \label{eq: asp of branching mechanism}
    \int_{(1,\infty)}y^{1+\beta +\delta}~\Big|\pi(dy)-\frac{\eta~dy}{\Gamma(-1-\beta)y^{2+\beta}}\Big| <\infty,
  \end{align}
	for some $\delta > 0$.
\end{asp}
We will show in Subsection \ref{sec: branching mechanism} that if Assumption \ref{asp: branching mechanism} holds, then $\eta$ and $\beta$ are uniquely determined by the L\'evy measure $\pi$.
In the reminder of the paper, we will always use $\eta$ and $\beta$ to denote the constants in Assumption  \ref{asp: branching mechanism}.

@@

Denote by $\mathcal B(\mathbb R^d, \mathbb R)$ the space of all $\mathbb R$-valued Borel functions on $\mathbb R^d$.
Denote by $\mathcal B(\mathbb R^d, \mathbb R_+)$ the space of all $\mathbb R_+$-valued Borel functions on $\mathbb R^d$.
We use  $(P_t)_{t\geq 0}$ to denote the transition semigroup of $\xi$.	
Define
\begin{align}
  \label{eq: meansemigroup}
  P^{\alpha}_t f(x)
  :=
  e^{\alpha t} P_t f(x) 
  = \Pi_x [e^{\alpha t}f(\xi_t)]
  , \quad x\in \mathbb R^d,t\geq 0, f\in \mathcal B(\mathbb R^d, \mathbb R_+).
\end{align}
It is known that, see \cite[Proposition 2.27]{Li2011Measure-valued} for example, $(P^\alpha_t)_{t\geq 0}$ is the \emph{mean semigroup} of $X$, in the sense that
\begin{equation}
  \mathbb{P}_{\mu}[\langle f, X_t \rangle]
  = \langle P^\alpha_t f, \mu \rangle,
  \quad t\geq 0, f\in \mathcal B(\mathbb R^d, \mathbb R_+), \mu \in \mathcal M(\mathbb R^d).
\end{equation}
The limiting behavior of the super-OU process is closely related to the asymptotic property of this mean semigroup $(P^\alpha_t)_{t\geq 0}$, and therefore, to the property of the OU semigroup $(P_t)_{t\geq 0}$.

We now recall some basic spectral properties of $(P_t)_{t\geq 0}$ from \cite{MetafunePallaraPriola2002Spectrum}.
It is known that the OU process $\xi$ has an invariant density
\begin{align}
\label{invariantdensity}
  \varphi(x)dx
  :=\Big (\frac{b}{\pi \sigma^2}\Big )^{d/2}\exp \Big(-\frac{b}{\sigma^2}|x|^2 \Big)dx,
  \quad x\in \mathbb R^d.
\end{align}
Let $L^2(\varphi):= \left\{ h  \in \mathcal B(\mathbb R^d, \mathbb R): \int_{\mathbb R^d} |h(x)|^2 \varphi(x) dx < \infty \right\}$.
Then $L^2(\varphi)$ is a Hilbert space with inner product
\begin{align}
  \langle f_1, f_2 \rangle_{\varphi}
  := \int_{\mathbb R^d}f_1(x)f_2(x)\varphi(x) dx, \quad f_1,f_2 \in L^2(\varphi).
\end{align}
Let $\mathbb Z_+ := \mathbb N\cup\{0\}$.
For each $p = (p_k)_{k = 1}^d \in \mathbb{Z}_+^{d}$, write $|p|:=\sum_{k=1}^d p_k$, $p!:= \prod_{k= 1}^d p_k!$ and $\partial x^p:= \prod_{k = 1}^d\partial x_k^{p_k}$.
The \emph{Hermite polynomials} are defined by
\begin{align}
  H_p(x)
  :=(-1)^{|p|}\exp(|x|^2) \frac{\partial ^{|p|}}{\partial x^p} \exp(-|x|^2) 
  , \quad x\in \mathbb R^d, p \in \mathbb{Z}_+^{d}.
\end{align}
It is known that $(P_t)_{t\geq 0}$ is a strongly continuous semigroup in $L^2(\varphi)$ and its generator $L$ has discrete spectrum $\sigma(L)= \{-bk: k \in \mathbb Z_+\}$.
For each $k \in \mathbb Z_+$, denote by $\mathcal{A}_k$ the eigenspace corresponding to the eigenvalue $-bk$, then
$
\mathcal{A}_k
= \operatorname{Span} \{\phi_p : p\in \mathbb Z_+^d, |p|=k\},
$
where
\begin{align}
  \label{eigenfunction}
  \phi_p(x)
  := \frac{1}{\sqrt{ p! 2^{|p|} }} H_p \Big(\frac{ \sqrt{b} }{\sigma}x \Big)
  , \quad x\in \mathbb R^d, p\in \mathbb Z_+^d.
\end{align}
In other words,
\begin{equation}
  P_t\phi_p(x)
  =e^{-b|p|t}\phi_p(x),
  \quad t\geq 0, x\in \mathbb R^d, p\in \mathbb Z_+^d.
\end{equation}
Moreover, $\{\phi_p: p \in \mathbb Z_+^d\}$ forms a complete orthonormal basis of $L^2(\varphi)$.
Thus for each $f\in L^2(\varphi)$, we have
\begin{equation}
\label{semicomp1}
f
=\sum_{k=0}^{\infty}\sum_{p\in \mathbb Z_+^d:|p|=k}\langle f, \phi_p \rangle_{\varphi} \phi_p
, \quad \text{in~} L^2(\varphi).
\end{equation}
For each function $f\in L^2(\varphi)$, denote by
\begin{equation}
  \kappa_f
  :=\inf \left \{k\geq 0: \exists ~ p\in \mathbb Z_+^d ,{\rm ~s.t.~}|p|=k {\rm ~and~}  \langle f, \phi_p \rangle_{\varphi}\neq 0\right \},
\end{equation}
the order of the function $f$ in the spectral decomposition \eqref{semicomp1}.
$\kappa_f$ corresponds to the lowest non-trivial frequency in the eigen-expansion of $f$.
Note that $ \kappa_f\geq 0$ and that, if $f\in L^2(\varphi)$ is non-trivial, then $\kappa_f<\infty$.
In particular, the order of any constant non-zero function is zero.
Denote by $\mathcal P$ the class of functions of polynomial growth on $\mathbb R^d$, i.e.,
\begin{equation}
  \label{eq: polynomial growth function}
  \mathcal{P}
  :=\big \{f\in \mathscr B(\mathbb R^d, \mathbb R):\exists~ C>0, n \in \mathbb Z_+, {\rm such  \, \, that \, } \forall x\in \mathbb R^d,~ |f(x)|\leq C(1+|x|)^n\big\}.
\end{equation}
It is clear that $\mathcal{P} \subset L^2(\varphi)$.
In this paper, we are mainly interested in the asymptotic behavior of $\langle f, X_t\rangle$ for non-trivial $f\in \mathcal P$.
The asymptotic behavior varies with the following three regimes:
(1) the small branching rate regime where $\alpha\beta<\kappa_fb(1+\beta)$; 
(2) the critical branching rate regime where $\alpha\beta=\kappa_fb(1+\beta)$; and
(3) the large branching rate regime where $\alpha\beta>\kappa_fb(1+\beta)$.

% *** Large branching rate regime
\subsubsection{Large branching rate regime}
Denote by $\mathcal M_c(\mathbb R^d)$ the space of all finite Borel measures of compact support on $\mathbb R^d$.
For each $p\in \mathbb{Z}_+^d$, define
\[
  H_t^p
  := e^{-(\alpha-|p|b)t}\langle\phi_p,X_t\rangle,
  \quad t\geq 0.
\]
If $\alpha\beta>|p|b(1+\beta)$, then for all $\gamma\in (0, \beta)$ and $\mu\in \mathcal M_c(\mathbb R^d)$, we will prove in Lemma \ref{lemma26} that $(H_t^p)_{t\geq 0}$ is a $\mathbb{P}_{\mu}$-martingale bounded in $L^{1+\gamma}(\mathbb{P}_{\mu})$.
Thus the limit $H^p_{\infty}:=\lim_{t\rightarrow \infty}H_t^p$ exists $\mathbb{P}_{\mu}$-almost surely and in $L^{1+\gamma}(\mathbb{P}_{\mu})$.
\begin{thm}
  \label{thm: law of large number}
  If $f \in \mathcal{P}$ satisfies $\alpha\beta>\kappa_fb(1+\beta)$, then for all $\gamma\in (0, \beta)$ and  $\mu\in \mathcal M_c(\mathbb R^d)$,
  \[
    e^{-(\alpha-\kappa_fb)t}\langle f, X_t\rangle
    \xrightarrow[t\to \infty]{}\sum_{p\in \mathbb Z_+^d:|p|=\kappa_f}\langle f, \phi_p\rangle_{\varphi} H_{\infty}^p
    \quad in~ L^{1+\gamma}(\mathbb{P}_{\mu}).
  \]
  Moreover, if $f$ is twice differentiable and all its second order partial derivatives are in $\mathcal{P}$, then we also have almost sure convergence.
\end{thm}
\begin{rem}
  If $f\in \mathcal B(\mathbb R^d, \mathbb R_+)$ is non-trivial and  bounded, then $\kappa_f=0$.
  Hence, Theorem \ref{thm: law of large number} says that for any $\gamma\in (0, \beta)$ and  $\mu\in \mathcal M_c(\mathbb R^d)$, as $t\rightarrow \infty$,
  \[
    e^{-\alpha t}\langle f, X_t\rangle
    \rightarrow \langle f, \varphi\rangle H_{\infty}
    \quad \mbox{in } L^{1+\gamma}(\mathbb{P}_{\mu}).
  \]
  Moreover, if $f$ is twice differentiable and all its second order partial derivatives are in $\mathcal{P}$, then we also have a.\/s.\ convergence.
  However, to get  a.\/s.\ convergence for bounded non-negative 
  Lebesgue-a.\/e.\/\ continuous functions $f$, we do not need $f$ to be twice differentiable. See \cite[Theorem 2.13 \& Example 8.1]{ChenRenYang2019Skeleton} and \cite[Theorem 1.2 \& Example 4.1]{EckhoffKyprianouWinkel2015Spines}.
\end{rem}
\begin{rem}
  For general $f\in \mathcal{P}$ with $\alpha\beta>\kappa_fb(1+\beta)$, Theorem \ref{thm: law of large number} extends the strong laws of large numbers of \cite{ChenRenYang2019Skeleton, EckhoffKyprianouWinkel2015Spines} in which the first order asymptotic ($\kappa_f=0$) was identified.
\end{rem}

>>>>

% ** Introduction

@@

% *** Motivation

@@

\label{subsec:M}

@@

  - \alpha z + \rho z^2 + \int_{(0,\infty)} (e^{-zy} - 1 + zy)~\pi(dy)

@@

where $\alpha > 0 $, $\rho \geq0$ and $\pi$ is a measure on $(0,\infty)$ with $\int_{(0,\infty)}(y\wedge y^2)~\pi(dy)< \infty$.
% $\psi$ is referred to as a branching mechanism and $\pi$ is referred to as the L\'evy measure of $\psi$.
$\psi$ is referred to as the branching mechanism and $\pi$ is referred to as the L\'evy measure of $\psi$.

@@

% For each $f,g\in \mathcal B(\mathbb R^d, \mathbb R)$ and $\mu \in \mathcal M(\mathbb R^d)$, write $\langle f,\mu\rangle = \int f(x)\mu(dx)$ and $\langle f, g\rangle = \int f(x)g(x) dx$ whenever the integrals make sense.
For each $f,g\in \mathcal B(\mathbb R^d, \mathbb R)$ and $\mu \in \mathcal M(\mathbb R^d)$, write $\mu(f)=\langle f,\mu\rangle = \int f(x)\mu(dx)$ and $\langle f, g\rangle = \int f(x)g(x) dx$ whenever the integrals make sense.

@@

% where $(t,x) \mapsto V_tf(x)$ is the unique locally bounded positive solution to the equation
where $(t,x) \mapsto V_tf(x)$ is the unique locally bounded non-negative solution to the equation

@@

% The strong law of large numbers for the super-OU process $X$ can be stated as follows:
The strong law of large numbers for the super-OU process $X$ introduced at the beginning can be stated as follows:
Under some conditions on the branching mechanism $\psi$ (these conditions are satisfied under our Assumptions 1 and 2 below), there exists an $\Omega_0$ of $\mathbb{P}_\mu$-full probability for every $\mu\in\mathcal M(\mathbb R^d)$ such that on $\Omega_0$, for every Lebesgue-a.\/e. continuous bounded non-negative function $f$ on $\mathbb R^d$, we have $\lim_{t\to\infty} e^{-\alpha t} \langle f, X_t\rangle =H_\infty\langle f, \varphi\rangle $, where $H_\infty$ is the limit of the martingale $e^{-\alpha t}\langle 1,X_t\rangle$ and $\varphi$ is the invariant density of the OU process $\xi$ defined in \eqref{invariantdensity} below.
See \cite[Theorem 2.13 \& Example 8.1]{ChenRenYang2019Skeleton} and \cite[Theorem 1.2 \& Example 4.1]{EckhoffKyprianouWinkel2015Spines}.

% In this paper, we will establish some spatial central limit theorems (CLT) for the super-OU process $X$ above.
In this paper, we will establish some spatial central limit theorems (CLTs) for the super-OU process $X$ above.
% Our key assumption is that the branching mechanism $\psi(z)$ takes the form of $-\alpha z +\rho z^2+ \psi_1(z)$, where $\psi_1$ is close to $\eta z^{1+\beta}$ with $\eta>0$ and $\beta\in (0, 1)$, in some sense.
Our key assumption is that the branching mechanism $\psi(z)$ takes the form of $-\alpha z +\rho z^2+ \psi_1(z)$, where $\psi_1$ is close to $\eta z^{1+\beta}$ with $\eta>0$ and $\beta\in (0, 1)$, in some sense.
% We want to find $(F_t)_{t\geq 0}$ and $(G_t)_{t\geq 0}$ such that $ (\langle f, X_t \rangle -G_t)/F_t $ converges weakly to some non-degenerate random variable as $t\rightarrow\infty$, for a large class of functions $f$.
And the general goal is to find $(F_t)_{t\geq 0}$ and $(G_t)_{t\geq 0}$ such that $ (\langle f, X_t \rangle -G_t)/F_t $ converges weakly to some non-degenerate random variable as $t\rightarrow\infty$, for a large class of functions $f$.
% DELETED
% It turns out that the statements of the CLT's are different in three regimes depending on the sign of $\alpha\beta-\kappa_f b (1+\beta)$: 
% the small branching rate regime $\alpha\beta< \kappa_f b(1+\beta)$; 
% the critical branching rate regime $\alpha\beta = \kappa_f b(1+\beta)$; 
% and the large branching rate regime $\alpha\beta > \kappa_f b(1+\beta)$. 
% Here, $\kappa_f$ is the order of $f$ in the spectral decomposition in terms of the spectrum of $L$.
% END DELETED
Note that, in the setting of this paper, $\langle f,X_t\rangle$ typically have infinite second moment.

% There are many papers studying central limit theorems for branching processes, branching diffusions and superprocesses under the second moment condition.
There are many papers studying CLTs for branching processes, branching diffusions and superprocesses under the second moment condition.
See \cite{Heyde1970A-rate, HeydeBrown1871An-invariance, HeydeLeslie1971Improved} for supercritical Galton-Watson processes (GW processes), \cite{KestenStigum1966Additional,KestenStigum1966A-limit} for supercritical multi-type GW processes, \cite{Athreya1969Limit,Athreya1969LimitB,Athreya1971Some} for supercritical multi-type continuous time branching processes and \cite{AsmussenHering1983Branching} for general supercritical branching Markov processes under certain conditions.
% Some spatial central limit theorems for   supercritical branching OU processes with binary branching mechanism were proved in \cite{AdamczakMilos2015CLT} and some spatial central limit theorems for supercritical super-OU processes with branching mechanisms satisfying a fourth moment condition were proved in \cite{Milos2012Spatial}.
Some spatial CLTs for   supercritical branching OU processes with binary branching mechanism were proved in \cite{AdamczakMilos2015CLT} and some spatial CLTs for supercritical super-OU processes with branching mechanisms satisfying a fourth moment condition were proved in \cite{Milos2012Spatial}.
These two papers made connections between the CLTs and  the branching rate regimes.
Some spatial central limit theorems for supercritical super-OU  processes with branching mechanisms satisfying only a second moment condition were established in \cite{RenSongZhang2014Central}.
Moreover, compared with the results of \cite{AdamczakMilos2015CLT,Milos2012Spatial}, the limit distributions in \cite{RenSongZhang2014Central} are non-degenerate.
Since then, a series of spatial central limit theorems for a large class of general supercritical branching Markov processes and superprocesses with spatially dependent branching mechanisms were proved in \cite{RenSongZhang2014CentralB,RenSongZhang2015Central,RenSongZhang2017Central}.
The functional version of the central limit theorems was established in \cite{Janson2004Functional} for supercritical multitype branching processes, and  in \cite{RenSongZhang2017Functional} for supercritical superprocesses.

@@

% *** Main results
\subsection{Main results}
\label{sec:I:R}
% **** Assumption 1
We will always assume that the following assumption holds.
\begin{asp}
  \label{asp: Greys condition}
  The branching mechanism satisfies Grey's condition, i.e., there exists $z' > 0$ such that $\psi(z) > 0$ for all $z>z'$ and  $\int_{z'}^\infty \psi(z)^{-1}dz < \infty$.
\end{asp}
For each $\mu \in \mathcal M(\mathbb R^d)$, write $\|\mu\| = \langle 1, \mu\rangle$.
It is known (see \cite[Theorems 12.5 \& 12.7]{Kyprianou2014Fluctuations} for example) that, under Assumption \ref{asp: Greys condition}, the \emph{extinction event} $D :=\{\exists t\geq 0,~\text{s.t.}~ \|X_t\| =0 \}$ has positive probability, with respect to $\mathbb P_\mu$ for each  $\mu \in \mathcal M(\mathbb R^d)$.
In fact, $ \mathbb{P}_{\mu} (D) = e^{-\bar v \|\mu\|}$ for each $\mu\in \mathcal M(\mathbb R^d)$ where $ \bar v := \sup\{\lambda \geq 0: \psi(\lambda) = 0\} \in (0,\infty) $ is the largest root of $\psi$.

% **** Assumption 2
Denote by $\Gamma$ the gamma function.
For any $\sigma$-finite signed measure $\mu$, we use $|\mu|$ to denote the total variation measure of $\mu$.
In this paper, we will also assume the following:
\begin{asp}
  \label{asp: branching mechanism}
  There exist constants $\eta > 0$ and $\beta \in (0,1)$ such that
  \begin{align}
    \label{eq: asp of branching mechanism}
    \int_{(1,\infty)}y^{1+\beta +\delta}~\Big|\pi(dy)-\frac{\eta~dy}{\Gamma(-1-\beta)y^{2+\beta}}\Big| <\infty,
  \end{align}
	for some $\delta > 0$.
\end{asp}
We will show in Subsection \ref{sec: branching mechanism} that if Assumption \ref{asp: branching mechanism} holds, then $\eta$ and $\beta$ are uniquely determined by the L\'evy measure $\pi$.
In the reminder of the paper, we will always use $\eta$ and $\beta$ to denote the constants in Assumption  \ref{asp: branching mechanism}.

@@

% **** Mean semigroup
Denote by $\mathcal B(\mathbb R^d, \mathbb R)$ the space of all $\mathbb R$-valued Borel functions on $\mathbb R^d$.
Denote by $\mathcal B(\mathbb R^d, \mathbb R_+)$ the space of all $\mathbb R_+$-valued Borel functions on $\mathbb R^d$.
We use  $(P_t)_{t\geq 0}$ to denote the transition semigroup of $\xi$.	
Define
\begin{align}
  \label{eq: meansemigroup}
  P^{\alpha}_t f(x)
  :=
  e^{\alpha t} P_t f(x) 
  = \Pi_x [e^{\alpha t}f(\xi_t)]
  , \quad x\in \mathbb R^d,t\geq 0, f\in \mathcal B(\mathbb R^d, \mathbb R_+).
\end{align}
It is known that, see \cite[Proposition 2.27]{Li2011Measure-valued} for example, $(P^\alpha_t)_{t\geq 0}$ is the \emph{mean semigroup} of $X$, in the sense that
\begin{equation}
  \mathbb{P}_{\mu}[\langle f, X_t \rangle]
  = \langle P^\alpha_t f, \mu \rangle,
  \quad t\geq 0, f\in \mathcal B(\mathbb R^d, \mathbb R_+), \mu \in \mathcal M(\mathbb R^d).
\end{equation}
The limiting behavior of the super-OU process is closely related to the asymptotic property of this mean semigroup $(P^\alpha_t)_{t\geq 0}$, and therefore, to the property of the OU semigroup $(P_t)_{t\geq 0}$.

% **** Spectral Property
We now recall some basic spectral properties of $(P_t)_{t\geq 0}$ from \cite{MetafunePallaraPriola2002Spectrum}.
It is known that the OU process $\xi$ has an invariant density
\begin{align}
\label{invariantdensity}
  \varphi(x)dx
  :=\Big (\frac{b}{\pi \sigma^2}\Big )^{d/2}\exp \Big(-\frac{b}{\sigma^2}|x|^2 \Big)dx,
  \quad x\in \mathbb R^d.
\end{align}
Let $L^2(\varphi):= \left\{ h  \in \mathcal B(\mathbb R^d, \mathbb R): \int_{\mathbb R^d} |h(x)|^2 \varphi(x) dx < \infty \right\}$.
Then $L^2(\varphi)$ is a Hilbert space with inner product
\begin{align}
  \langle f_1, f_2 \rangle_{\varphi}
  := \int_{\mathbb R^d}f_1(x)f_2(x)\varphi(x) dx, \quad f_1,f_2 \in L^2(\varphi).
\end{align}
Let $\mathbb Z_+ := \mathbb N\cup\{0\}$.
% For each $p = (p_k)_{k = 1}^d \in \mathbb{Z}_+^{d}$, write $|p|:=\sum_{k=1}^d p_k$, $p!:= \prod_{k= 1}^d p_k!$ and $\partial x^p:= \prod_{k = 1}^d\partial x_k^{p_k}$.
For each $p = (p_k)_{k = 1}^d \in \mathbb{Z}_+^{d}$, write $|p|:=\sum_{k=1}^d p_k$, $p!:= \prod_{k= 1}^d p_k!$ and $\partial_p:= \prod_{k = 1}^d(\partial/\partial x_k^{p_k})$.
The \emph{Hermite polynomials} are defined by
\begin{align}
  H_p(x)
  % :=(-1)^{|p|}\exp(|x|^2) \frac{\partial ^{|p|}}{\partial x^p} \exp(-|x|^2) 
  :=(-1)^{|p|}\exp(|x|^2) \partial_p \exp(-|x|^2) 
  , \quad x\in \mathbb R^d, p \in \mathbb{Z}_+^{d}.
\end{align}
It is known that $(P_t)_{t\geq 0}$ is a strongly continuous semigroup in $L^2(\varphi)$ and its generator $L$ has discrete spectrum $\sigma(L)= \{-bk: k \in \mathbb Z_+\}$.
For each $k \in \mathbb Z_+$, denote by $\mathcal{A}_k$ the eigenspace corresponding to the eigenvalue $-bk$, then
$
\mathcal{A}_k
= \operatorname{Span} \{\phi_p : p\in \mathbb Z_+^d, |p|=k\},
$
where
\begin{align}
  \label{eigenfunction}
  \phi_p(x)
  := \frac{1}{\sqrt{ p! 2^{|p|} }} H_p \Big(\frac{ \sqrt{b} }{\sigma}x \Big)
  , \quad x\in \mathbb R^d, p\in \mathbb Z_+^d.
\end{align}
In other words,
\begin{equation}
  P_t\phi_p(x)
  =e^{-b|p|t}\phi_p(x),
  \quad t\geq 0, x\in \mathbb R^d, p\in \mathbb Z_+^d.
\end{equation}
Moreover, $\{\phi_p: p \in \mathbb Z_+^d\}$ forms a complete orthonormal basis of $L^2(\varphi)$.
Thus for each $f\in L^2(\varphi)$, we have
\begin{equation}
\label{semicomp1}
f
=\sum_{k=0}^{\infty}\sum_{p\in \mathbb Z_+^d:|p|=k}\langle f, \phi_p \rangle_{\varphi} \phi_p
, \quad \text{in~} L^2(\varphi).
\end{equation}
For each function $f\in L^2(\varphi)$, denote by
\begin{equation}
  \kappa_f
  :=\inf \left \{k\geq 0: \exists ~ p\in \mathbb Z_+^d ,{\rm ~s.t.~}|p|=k {\rm ~and~}  \langle f, \phi_p \rangle_{\varphi}\neq 0\right \},
\end{equation}
the order of the function $f$ in the spectral decomposition \eqref{semicomp1}.
$\kappa_f$ corresponds to the lowest non-trivial frequency in the eigen-expansion of $f$.
Note that $ \kappa_f\geq 0$ and that, if $f\in L^2(\varphi)$ is non-trivial, then $\kappa_f<\infty$.
In particular, the order of any constant non-zero function is zero.

% **** Martingale limits
% MOVED from below
  Denote by $\mathcal M_c(\mathbb R^d)$ the space of all finite Borel measures of compact support on $\mathbb R^d$.
  For each $p\in \mathbb{Z}_+^d$, define
  \[
    H_t^p
    := e^{-(\alpha-|p|b)t}\langle\phi_p,X_t\rangle,
    \quad t\geq 0.
  \]
  If $\alpha\beta>|p|b(1+\beta)$, then for all $\gamma\in (0, \beta)$ and $\mu\in \mathcal M_c(\mathbb R^d)$, we will prove in Lemma \ref{lem:M:L:ML} that $(H_t^p)_{t\geq 0}$ is a $\mathbb{P}_{\mu}$-martingale bounded in $L^{1+\gamma}(\mathbb{P}_{\mu})$.
  Thus the limit $H^p_{\infty}:=\lim_{t\rightarrow \infty}H_t^p$ exists $\mathbb{P}_{\mu}$-almost surely and in $L^{1+\gamma}(\mathbb{P}_{\mu})$.
% END MOVED

% **** LLNs
% ADDED
We first present a law of large numbers for our model which extends the strong laws of large numbers of \cite{ChenRenYang2019Skeleton, EckhoffKyprianouWinkel2015Spines} in which the first order asymptotic ($\kappa_f=0$) was identified.
% END ADDED
Denote by $\mathcal P$ the class of functions of polynomial growth on $\mathbb R^d$, i.e.,
\begin{equation}
  \label{eq: polynomial growth function}
  \mathcal{P}
  % :=\big \{f\in \mathscr B(\mathbb R^d, \mathbb R):\exists~ C>0, n \in \mathbb Z_+, {\rm such  \, \, that \, } \forall x\in \mathbb R^d,~ |f(x)|\leq C(1+|x|)^n\big\}.
  :=\big \{f\in \mathcal B(\mathbb R^d, \mathbb R):\exists~ C>0, n \in \mathbb Z_+, {\rm such  \, \, that \, } \forall x\in \mathbb R^d,~ |f(x)|\leq C(1+|x|)^n\big\}.
\end{equation}
It is clear that $\mathcal{P} \subset L^2(\varphi)$.
% MOVED FROM BELOW
\begin{thm}
  \label{thm: law of large number}
  If $f \in \mathcal{P}$ satisfies $\alpha\beta>\kappa_fb(1+\beta)$, then for all $\gamma\in (0, \beta)$ and  $\mu\in \mathcal M_c(\mathbb R^d)$,
  \[
    e^{-(\alpha-\kappa_fb)t}\langle f, X_t\rangle
    \xrightarrow[t\to \infty]{}\sum_{p\in \mathbb Z_+^d:|p|=\kappa_f}\langle f, \phi_p\rangle_{\varphi} H_{\infty}^p
    \quad in~ L^{1+\gamma}(\mathbb{P}_{\mu}).
  \]
  Moreover, if $f$ is twice differentiable and all its second order partial derivatives are in $\mathcal{P}$, then we also have almost sure convergence.
\end{thm}
If $f\in \mathcal B(\mathbb R^d, \mathbb R_+)$ is non-trivial and  bounded, then $\kappa_f=0$.
Hence, Theorem \ref{thm: law of large number} says that for any $\gamma\in (0, \beta)$ and  $\mu\in \mathcal M_c(\mathbb R^d)$, as $t\rightarrow \infty$,
\[
  e^{-\alpha t}\langle f, X_t\rangle
  \rightarrow \langle f, \varphi\rangle H_{\infty}
  \quad \mbox{in } L^{1+\gamma}(\mathbb{P}_{\mu}).
\]
Moreover, if $f$ is twice differentiable and all its second order partial derivatives are in $\mathcal{P}$, then we also have a.\/s.\ convergence.
However, to get  a.\/s.\ convergence for bounded non-negative 
Lebesgue-a.\/e.\/\ continuous functions $f$, we do not need $f$ to be twice differentiable. See \cite[Theorem 2.13 \& Example 8.1]{ChenRenYang2019Skeleton} and \cite[Theorem 1.2 \& Example 4.1]{EckhoffKyprianouWinkel2015Spines}.
% DELETED
%\begin{rem}
%  For general $f\in \mathcal{P}$ with $\alpha\beta>\kappa_fb(1+\beta)$, Theorem \ref{thm: law of large number} extends the strong laws of large numbers of \cite{ChenRenYang2019Skeleton, EckhoffKyprianouWinkel2015Spines} in which the first order asymptotic ($\kappa_f=0$) was identified.
%\end{rem}
% END DELETED
% END MOVED FROM BELOW
